\chapter*{Preface}

\section*{Supplementary Material}


Instructions and the relevant code needed to reproduce this thesis can be found in the \href{https://github.com/LGraz/MasterThesis-Code}{\color{blue}{GitHub repository}} and to use our results we recommend the provided \href{https://github.com/LGraz/CorrectTimeSeries }{\color{blue}{R-package}}.

More information is given in the appendix~\ref{app:reproducibility}.  


\section*{Acknowledgements}
First, I wish to express my sincere gratitude to my supervisor Prof. Dr. Nicolai Meinshausen who took the responsibility for my work and happily took the time to discuss conceptual and guiding questions and to inspire me with new ideas. 

It is necessary to highlight that without Gregor Perich this project would not have been possible. His high personal commitment, reliability as well as the weekly instructive supervision meetings were, without question, essential for this work. 

It was a real pleasure for me to be part of the \textit{Crop Science} group for this time. Enjoying everyday company, a two-day excursion, and harvesting wheat together have made this time truly remarkable. In particular, I would like to thank Prof. Dr. Achim Walter, who supported this collaboration at its core.  

Last but not least, I would like to express my gratitude to the \textit{Seminar for Statistics}, which created the framework conditions for this work and did everything to help me with conceptional and administrative questions. I should also mention the computing resources provided by them, without which my computations would not have been feasible.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "MasterThesisSfS"
%%% End: 
