\chapter{Results}

\section{Goodness of Fit for Selected Interpolation Methods}{
	\label{sec:results_itpl}
	Table \ref{tab:cv-statistics_itpl-methods} benchmarks the selected\footnote{\label{note:selected-IM} For the discussion which methods have been selected c.f. section \ref{sec:itpl_preselection}.} interpolation methods (on $P^{SCL45}$) with respect to various score functions. The score functions take the absolute values of the LOOCV residuals and summarize them in a number (the smaller, the better). For each of the 5 selected interpolation methods, we consider the basic and the robustified (see section \ref{sec:loess_robustify}) version.

	\begin{table}[h]
		\begin{center}
			\caption{Comparing the goodness of fit for selected%\footnoteref{note:selected-IM} % not working :-(
			interpolation methods (on $P^{SCL45}$) measured with the score functions (which take the LOOCV residuals as input) listed in the left column. $q_X$ denotes here the $X\%$ quantile.}
			\scriptsize
			\input{tex/chapters/misc/table_cv-statistics_itpl-methods}
			\normalsize
			\label{tab:cv-statistics_itpl-methods}
		\end{center}
	\end{table}
	
	DL is the best among both robustified and non-robustified with respect to most of the score functions used (all except q95) and is especially superior to the other parametric approach, which is Fourier interpolation. Especially the robust Fourier interpolation performs poorly. The LOESS dominates (i.e. is superior on every score function) all other non-prametric methods, but is closely followed by the SS. The BSPL, on the other hand, is the worst non-parametric method tested here. 
}


\section{XXX (Robustification and) NDVI-Correction} \label{sec:results_ndvi_corr}
\todo[inline]{defiition of RYEA, it is not an accuracy but an error}
	The RYEA for the 28 (in section \ref{sec:corr_itpl_stat}) chosen interpolation strategies is given in table \ref{tab:methods_vs_yieldprediction_relative}. 

	Robustification in the interpolation strategies, does not improve the quality of the fit (measured via the RYEA) in most cases. 

	Abgesehen von einer ausnahme sind SS (rob) besser als DL(rob) hinsichtlich der RYEA.

	Die Interpolations strategie, welche die niedrigste RYEA hervorbringt ist OLS-SCL mit SS. Da die OLS-SCL modelle eine sehr gute interpretierbarkeit haben präsentieren wir im folgenden auch die Regressionsgleichungen. Der korrigierten NDVI wird berechnet anhand von EQUATION, wo $\mathbbm 1_{{SCL = 2}}$ gleich eins ist, falls die aktuelle beobachtung der SCL Klasse 2 entspricht und null andernfalls.\footnote{$\mathbbm 1$ wird in der Mathematik auch als indikatorfunktion oder charakteristische funktion bezeichent.} Die geschätzen absoluten residuen erhalten wir hingegen durch: EQUATION  
	Für den R-output der summary funktion der beiden modelle verweisen wir auf den Anhang 

	Für die conflusion:
	Wir wollen die in kapitel \ref{sec:corr} eingeführe interpolationsstrategie nochmals zusammenfassen: Wir schätzen den echten NDVI anhand von SS via LOOCV, dann erhalten wir den corrigierten NDVI mit dem OLS-SCL modell. Anschließend schätzen wir den absoluten fehler mit dem OLS-SCL modell und erhalten somit gewichte, welche die zuverlässigkeit der korrigerten NDVI wiederspiegeln sollen. Zum schluss führen wir eine gewichtete interpolation mit SS durch.



\begin{table}
	\begin{center}
		\caption{RYEA. For the non-relative RMSE and the coefficient of determination (R\textsuperscript{2}) see table \ref{tab:methods_vs_yieldprediction} and \ref{tab:methods_vs_yieldprediction_r2}.}
		\small
		\input{tex/chapters/misc/table_methods_vs_yieldprediction_relative.tex}
		\label{tab:methods_vs_yieldprediction_relative}
		\normalsize
	\end{center}
\end{table}



\begin{equation}\label{eq:corr_lm}
	\begin{aligned}		
		{\operatorname{NDVI}}_{\text{corr}}  = &
		0.711 \operatorname{NDVI}_\text{observed}  
		+ \mathbbm 1_{{SCL = 2}} 0.215 
		+ \mathbbm 1_{{SCL = 3}} 0.237 
		+ \mathbbm 1_{{SCL = 4}} 0.210 \\ &
		+ \mathbbm 1_{{SCL = 5}} 0.116 
		+ \mathbbm 1_{{SCL = 6}} 0.162 
		+ \mathbbm 1_{{SCL = 7}} 0.327 
		+ \mathbbm 1_{{SCL = 8}} 0.474 \\ &
		+ \mathbbm 1_{{SCL = 9}} 0.575 
		+ \mathbbm 1_{{SCL = 10}} 0.306 
		+ \mathbbm 1_{{SCL = 11}} 0.512 
	\end{aligned}
\end{equation}
- strong upwards correction for SCL classes 8, 9 and 11 (correspond to `medium probability clouds', `high probability clouds' and `thin cirrus clouds').

\begin{equation}\label{eq:corr_lm_res}
	\begin{aligned}		
		\widehat{\operatorname{abs}}\left({\operatorname{NDVI}_\text{``true''} - \operatorname{NDVI}}_{\text{corr}}\right)  = &
		-0.133 \operatorname{NDVI}_\text{observed}  
		+ \mathbbm 1_{{SCL = 2}} 0.186 
		+ \mathbbm 1_{{SCL = 3}} 0.185 \\ &
		+ \mathbbm 1_{{SCL = 4}} 0.146 
		+ \mathbbm 1_{{SCL = 5}} 0.089 
		+ \mathbbm 1_{{SCL = 6}} 0.167 \\ &
		+ \mathbbm 1_{{SCL = 7}} 0.203 
		+ \mathbbm 1_{{SCL = 8}} 0.181  
		+ \mathbbm 1_{{SCL = 9}} 0.173 \\ &
		+ \mathbbm 1_{{SCL = 10}} 0.180 
		+ \mathbbm 1_{{SCL = 11}} 0.172 
	\end{aligned}
\end{equation}
- the higher the observed NDVI the lower the estimated absolute residual.
- estimated absolute residuals are the smallest for SCL classes 4 and 5.  
